Title here

### 2016-09-26 portfolio/text/ms.Rmd
Strip trailing white space
Editing of paper

A fundamental challenge for individuals who rely on the extraction of natural
resource for their livelihood is income variability.
Fishers, loggers, farmers, miners, and others
all face uncertainty in income driven by a mix of
environmental, biological, and economic (e.g. price) variability.
A key component underlying the sustainability 
of these industries is identifying both
regulatory actions and individual behaviour affect income variability 
and by extension the economic stability and 
sustainable use of natural resources.   
For commercial fisheries, recent work has identified that 
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship 
is embedded in a regulatory framework that restricts individual ability 
to unilaterally change fishing behaviour.  
Here we identify three primary axes by which fishers in Alaska, 
one of the world's most productive fishery regions, can 
potentially diversify fishing activities and understand how 
individual actions along each diversity axis affect 
both revenue and revenue variability. 
We construct a novel statistical model to simultaneously 
assess the value to individuals fishers of choosing to 
(1) own permits that allow fishers that target a diverse fish species assemblage, 
(2) diversify by owning multiple permits, or 
(3) catch a greater species within the constraints of a given permit.  
Thus, we present a unique approach to understanding important aspects of
diversity relevant to individual fishers' decisions. 
We show contrasting effects of diversity on 
revenue and revenue variability along the three axes. 
First, we show that fishing permits with 
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit 
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits);
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.

Thus, while there are clear benefits 
to some flavours of diversification, others don't do much.
Importantly, the most effective measure for 
increasing revenue and decreasing risk --- 
purchasing additional permits to participate in additional fisheries --- 
was also the most costly.

At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

-Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff
up here and added to it.[*Eric: I think we need a little here on the AK data
and why that's both important and generalizable to fisheries around the world.]

@@

*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
Fisheries in Alaska are the highest valued fisheries in N. America
with landings valued at XX $ billion annually. There are a number of fishing
strategies where both specialization and diversification are extremely profitable. There's a number of fishing permits that constrain the potential diversity of species landed (e.g. herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.* 
*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*

@@

-Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
To jointly model revenue and revenue variability as a function of
individual-level covariates, we use a variance function regression model, which
separately models change in the mean and variance as
a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
linear models or hierarchical linear models generally assume constant
variance [Ole: across what? covariates?], the variance function regression model allows variance to be
heteroskedastic. For application to economic data on individual fishing
portfolios, this type of model allows certain strategies and levels of
diversification to be more or less variable than others.

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$ 
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:
Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.

@@

Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels.

>>>>

# Abstract

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability.
Fishers, loggers, farmers, miners, and others
all face uncertainty in income driven by a mix of
environmental, biological, and economic variability.
A key component underlying the sustainability
of these industries is identifying both
how regulatory actions and individual behaviour affect income variability
and by extension the economic stability and
sustainable use of natural resources.
For commercial fisheries, recent work has identified that
targeting a diversity of species can reduce income variability.
However, any diversity-variability relationship
is embedded in a regulatory framework that restricts individual ability
to unilaterally change fishing behaviour.
Here we identify three primary axes by which fishers in Alaska,
one of the world's most productive fishery regions, can
potentially diversify fishing activities and
investigate how individual actions along each diversity axis affect
both revenue and revenue variability.
We construct a novel statistical model to simultaneously
assess the value to individual fishers of choosing to
(1) own permits that allow targeting a diverse fish species assemblage,
(2) diversify by owning multiple permits, or
(3) catch a greater species diversity within the constraints of a given permit.
We show contrasting effects of diversity on
revenue and revenue variability along the three axes.
First, we show that fishing permits with
higher species diversity reduces variability
(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
increase in mean effective species diversity).
Next, we show that fishing an additional permit
is associated with higher revenue
and lower variability (e.g.,
`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
expected standard deviation
for halibut plus another permit vs. single halibut permits).
Finally, we show that increasing species diversity within
the constraints of a permit set has a fishery-dependent effect on revenue and
is usually (`r g2_dens`% probability)
associated with increased uncertainty in revenue the following year.
Thus, while there are clear benefits
to some axes of diversification, others are less effective.
Importantly, the most effective measure for
increasing revenue and decreasing risk ---
purchasing additional permits to participate in additional fisheries ---
was also the most costly.

At the same time, Alaskan commercial fisheries
have been increasingly focussed on
single-permit (often salmon) fishing strategies,
with single-permit fishers earning
more than multiple-permit fishers
for the first time in 2013.
Our results suggest that the benefit of species diversity
for revenue variability
is mostly an across-permit effect, not
a benefit that can be easily harnessed by an individual permit holder.
Effective methods of diversification ---
purchasing additional permits or switching permits ---
are costly and often limited by regulations such as limited entry
programs and individual transferable quotas (ITQs).
Take home ...

@@

<!-- -Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff -->
<!-- up here and added to it.[*Eric: I think we need a little here on the AK data -->
<!-- and why that's both important and generalizable to fisheries around the world.] -->

@@

@@

To jointly model revenue and revenue variability as a function of
individual-level covariates, we use a variance function regression model, which
separately models change in the mean and variance as
a function of covariates [e.g. @davidian1987; @hall1989]. 
Whereas traditional linear models or 
hierarchical linear models generally assume
homoscedasticity --- that the residual variance 
is the same for all values of the predictor variables ---
the variance function regression model allows variance to be heteroskedastic. 
For application to economic data on individual fishing
portfolios, this type of model allows certain strategies and levels of
diversification to be more or less variable than others.

@@

\noindent
Each year $t$ of revenue for fisher $i$
is assigned a strategy indexed by $j$
(Supporting Information, Table \ref{tab:strategies})
and expected $\log(R_{i,t})$ is modelled as:
<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
<!-- potential confusion like using the term "expected log(R)".   -->
<!-- I also just don't like hats... but that's my problem. -->

@@

<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
